Privacy-Preserving Localization Using Private Linear-Combination Aggregation

### 2020-11-27 main.tex
Fixed up notation that was technically incorrect. "randomness" removed from function inputs when referring to function inputs. Similarly for the definition of the state vector.

The navigator state captures the position, and change in position ($\Delta$), and is defined as 
\begin{equation}
    \rvec{x} = 
    \begin{bmatrix}
        \rv{x} & \Delta\rv{x} & \rv{y} & \Delta\rv{y}
    \end{bmatrix}^\top\,. \label{eqn:state_def}
\end{equation}
A known process model is followed, which at time $k$ is given by
\begin{equation}
    \rvec{x}_k = \vec{f}(\rvec{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $\rvec{w}_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependent on sensor $i$ and given by
\begin{equation}
    \rv{z}_{k,i} = h_i(\rvec{x}_k)+\rv{v}_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_{k,i} \sim \mathcal{N}(0,r_i)$ and measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
defined as
\begin{equation}
    \begin{split}
        h_i(\rvec{x}) &= \left\lVert
        \begin{bmatrix}
            \rv{x} & \rv{y}
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(\rv{x}-s_{x,i})^2 + (\rv{y}-s_{y,i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn state estimates of $\rvec{x}_{k}$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$, measurement variances $r_i$ or their measurements $\rv{z}_{k,i}$ at any time $k$. We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. The additional safety to passengers provided by accurate aircraft localization, however, may be a goal of all those involved, justifying their cooperation.

@@

\subsection{Extended Information Filter}
The Extended Information Filter (EIF) \cite{maybeckStochasticModelsEstimation1982} is a reformulation of the Extended Kalman Filter (EKF), such that the update equations are simplified to sums of measurement information from each sensor. This requires the conversion of predicted state estimate $\rvec{x}_{k|k-1}$ and estimate covariance $\mat{P}_{k|k-1}$ to the information vector and matrix
\begin{equation}
    \rvec{y}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\rvec{x}_{k|k-1} \quad\textrm{and}\quad \mat{Y}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\,,
\end{equation}
respectively. In this form, the update equations for $n$ measurements at time $k$, given process and measurement models \eqref{eqn:process_model} and \eqref{eqn:measurement_model}, are given by
\begin{equation}
    \begin{split}
        &\rvec{y}_{k|k} = \rvec{y}_{k|k-1} + \\
        &\quad \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i (z_{k,i} - h_i(\rvec{x}_{k|k-1}) + \mat{H}_{k,i}\rvec{x}_{k|k-1})
    \end{split} \label{eqn:eif_info_vec_update}
\end{equation}
and
\begin{equation}
    \mat{Y}_{k|k} = \mat{Y}_{k|k-1} + \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i \mat{H}_{k,i}\,, \label{eqn:eif_info_mat_update}
\end{equation}
with Jacobians
\begin{equation}
    \mat{H}_{k,i} = \left.\frac{\partial h_i}{\partial \rvec{x}}\right|_{\rvec{x}_{k|k-1}} \label{eqn:measurement_jacobian}
\end{equation}
for sensors $1 \leq i \leq n$. The filter's prediction step can be computed by converting information vector $\rvec{y}_{k|k}$ and matrix $\mat{Y}_{k|k}$ back to state estimate and covariance 
\begin{equation}
    \rvec{x}_{k|k} = \mat{Y}_{k|k}^{-1}\rvec{y}_{k|k} \quad\textrm{and}\quad \mat{P}_{k|k} = \mat{Y}_{k|k}^{-1}\,,
\end{equation}
before using the normal EKF prediction equations
\begin{equation}
    \rvec{x}_{k+1|k} = \vec{f}(\rvec{x}_{k|k})
\end{equation}
and
\begin{equation}
    \mat{P}_{k+1|k} = \mat{F}_k\mat{P}_{k|k}\mat{F}_k^\top\,,
\end{equation}
with Jacobian
\begin{equation}
    \mat{F}_{k} = \left.\frac{\partial \vec{f}}{\partial \rvec{x}}\right|_{\rvec{x}_{k|k}}\,. \label{eqn:jacobian_f}
\end{equation}

@@

\subsection{Localization Measurement Model} \label{subsec:observation_model}
Recalling the state definition \eqref{eqn:state_def} and two-dimensional measurement model \eqref{eqn:measurement_model}, we can see that $h_i$ \textit{cannot} be rearranged to a linear combination of elements private to the navigator. In addition, the Jacobian of $h_i$ at $\rvec{x}_{k|k-1}$,
\begin{equation}
    \mat{H}_{k,i} = 
    \begin{bmatrix}
        \frac{\rv{x}_{k|k-1} - s_{x,i}}{\sqrt{(\rv{x}_{k|k-1} - s_{x,i})^2 + (\rv{y}_{k|k-1} - s_{y,i})^2}} \\
        \frac{\rv{y}_{k|k-1} - s_{y,i}}{\sqrt{(\rv{x}_{k|k-1} - s_{x,i})^2 + (\rv{y}_{k|k-1} - s_{y,i})^2}} 
    \end{bmatrix}\,,
\end{equation}
cannot be either. Therefore, we consider the modified measurement functions
\begin{equation}
    h'_i(\rvec{x}) = h_i(\rvec{x})^2\,. \label{eqn:modified_measurement_func}
\end{equation}
A measurement function in this form allows rearrangement to a linear combination of navigator information, and thus of the corresponding modified measurement vector $\rvec{i}'_{k,i}$ and measurement matrix $\mat{I}'_{k,i}$ as well. The modified measurement functions $h'_i$ can be written as
\begin{equation}
    \begin{split}
        h'_i(\rvec{x}) &= \left\lVert
        \begin{bmatrix}
            \rv{x} & \rv{y}
        \end{bmatrix}^\top - \vec{s}_i\right\rVert^2 \\
        &= (\rv{x} - s_{x,i})^2 + (\rv{y} - s_{y,i})^2 \\
        &= \rv{x}^2 + \rv{y}^2 -2s_{x,i}\rv{x} -2s_{y,i}\rv{y} +s_{x,i}^2 +s_{y,i}^2
    \end{split}\,,
\end{equation}
and the corresponding Jacobians $\mat{H}'_{k,i}$ as
\begin{equation}
    \mat{H}'_{k,i} = 
    \begin{bmatrix}
        2\rv{x}_{k|k-1} - 2s_{x,i} \\
        2\rv{y}_{k|k-1} - 2s_{y,i}
    \end{bmatrix}\,. \label{eqn:modified_jacobian}
\end{equation}
From the above we see that $h'_i(\rvec{x}_{k|k-1})$ and $\mat{H}'_{k,i}$ can be rearranged as linear combinations of the weights $\rv{x}_{k|k-1}^2$, $\rv{y}_{k|k-1}^2$, $\rv{x}_{k|k-1}$ and $\rv{y}_{k|k-1}$, which capture all the required information that is private to the navigator.

>>>>

The navigator state captures the position and change in position ($\Delta$), and is of the form 
\begin{equation}
    \vec{x} = 
    \begin{bmatrix}
        x & \Delta x & y & \Delta y
    \end{bmatrix}^\top\,. \label{eqn:state_def}
\end{equation}
A known process model is followed, in which the state at time $k$, $\rvec{x}_k$, is given by
\begin{equation}
    \rvec{x}_k = \vec{f}(\rvec{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $\rvec{w}_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependent on sensor $i$ and given by
\begin{equation}
    \rv{z}_{k,i} = h_i(\rvec{x}_k)+\rv{v}_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_{k,i} \sim \mathcal{N}(0,r_i)$. The measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
is defined as
\begin{equation}
    \begin{split}
        h_i(\vec{x}) &= \left\lVert
        \begin{bmatrix}
            x & y
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(x-s_{x,i})^2 + (y-s_{y,i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn state estimates of $\rvec{x}_{k}$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$, measurement variances $r_i$ or their measurements $\rv{z}_{k,i}$ at any time $k$. We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. The additional safety to passengers provided by accurate aircraft localization, however, may be a goal of all those involved, justifying their cooperation.

@@

\subsection{Extended Information Filter}
The Extended Information Filter (EIF) \cite{maybeckStochasticModelsEstimation1982} is a reformulation of the Extended Kalman Filter (EKF), such that the update equations are simplified to sums of measurement information from each sensor. This requires the conversion of predicted state estimate $\rvec{x}_{k|k-1}$ and estimate covariance $\mat{P}_{k|k-1}$ to the information vector and matrix
\begin{equation}
    \rvec{y}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\rvec{x}_{k|k-1} \quad\textrm{and}\quad \mat{Y}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\,,
\end{equation}
respectively. In this form, the update equations for $n$ measurements at time $k$, given process and measurement models \eqref{eqn:process_model} and \eqref{eqn:measurement_model}, are given by
\begin{equation}
    \begin{split}
        &\rvec{y}_{k|k} = \rvec{y}_{k|k-1} + \\
        &\quad \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i (z_{k,i} - h_i(\rvec{x}_{k|k-1}) + \mat{H}_{k,i}\rvec{x}_{k|k-1})
    \end{split} \label{eqn:eif_info_vec_update}
\end{equation}
and
\begin{equation}
    \mat{Y}_{k|k} = \mat{Y}_{k|k-1} + \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i \mat{H}_{k,i}\,, \label{eqn:eif_info_mat_update}
\end{equation}
with Jacobians
\begin{equation}
    \mat{H}_{k,i} = \left.\frac{\partial h_i}{\partial \vec{x}}\right|_{\rvec{x}_{k|k-1}} \label{eqn:measurement_jacobian}
\end{equation}
for sensors $1 \leq i \leq n$. The filter's prediction step can be computed by converting information vector $\rvec{y}_{k|k}$ and matrix $\mat{Y}_{k|k}$ back to state estimate and covariance 
\begin{equation}
    \rvec{x}_{k|k} = \mat{Y}_{k|k}^{-1}\rvec{y}_{k|k} \quad\textrm{and}\quad \mat{P}_{k|k} = \mat{Y}_{k|k}^{-1}\,,
\end{equation}
before using the normal EKF prediction equations
\begin{equation}
    \rvec{x}_{k+1|k} = \vec{f}(\rvec{x}_{k|k})
\end{equation}
and
\begin{equation}
    \mat{P}_{k+1|k} = \mat{F}_k\mat{P}_{k|k}\mat{F}_k^\top\,,
\end{equation}
with Jacobian
\begin{equation}
    \mat{F}_{k} = \left.\frac{\partial \vec{f}}{\partial \vec{x}}\right|_{\rvec{x}_{k|k}}\,. \label{eqn:jacobian_f}
\end{equation}

@@

\subsection{Localization Measurement Model} \label{subsec:observation_model}
Recalling the state definition \eqref{eqn:state_def} and two-dimensional measurement model \eqref{eqn:measurement_model}, we can see that $h_i$ \textit{cannot} be rearranged to a linear combination of elements private to the navigator. In addition, the Jacobian of $h_i$ at $\rvec{x}_{k|k-1}$,
\begin{equation}
    \mat{H}_{k,i} = 
    \begin{bmatrix}
        \frac{\rv{x}_{k|k-1} - s_{x,i}}{\sqrt{(\rv{x}_{k|k-1} - s_{x,i})^2 + (\rv{y}_{k|k-1} - s_{y,i})^2}} \\
        \frac{\rv{y}_{k|k-1} - s_{y,i}}{\sqrt{(\rv{x}_{k|k-1} - s_{x,i})^2 + (\rv{y}_{k|k-1} - s_{y,i})^2}} 
    \end{bmatrix}\,,
\end{equation}
cannot be either. Therefore, we consider the modified measurement functions
\begin{equation}
    h'_i(\vec{x}) = h_i(\vec{x})^2\,. \label{eqn:modified_measurement_func}
\end{equation}
A measurement function in this form allows rearrangement to a linear combination of navigator information, and thus of the corresponding modified measurement vector $\rvec{i}'_{k,i}$ and measurement matrix $\mat{I}'_{k,i}$ as well. The modified measurement functions $h'_i$ can be written as
\begin{equation}
    \begin{split}
        h'_i(\vec{x}) &= \left\lVert
        \begin{bmatrix}
            x & y
        \end{bmatrix}^\top - \vec{s}_i\right\rVert^2 \\
        &= (x - s_{x,i})^2 + (y - s_{y,i})^2 \\
        &= x^2 + y^2 -2s_{x,i}x -2s_{y,i}y +s_{x,i}^2 +s_{y,i}^2
    \end{split}\,,
\end{equation}
and the corresponding Jacobians $\mat{H}'_{k,i}$ as
\begin{equation}
    \mat{H}'_{k,i} = 
    \begin{bmatrix}
        2\rv{x}_{k|k-1} - 2s_{x,i} \\
        2\rv{y}_{k|k-1} - 2s_{y,i}
    \end{bmatrix}\,. \label{eqn:modified_jacobian}
\end{equation}
From the above we see that $h'_i(\rvec{x}_{k|k-1})$ and $\mat{H}'_{k,i}$ can be rearranged as linear combinations of the weights $\rv{x}_{k|k-1}^2$, $\rv{y}_{k|k-1}^2$, $\rv{x}_{k|k-1}$ and $\rv{y}_{k|k-1}$, which capture all the required information that is private to the navigator.
